---
title: "NPOR: Data preparation"
author: jellevanderwerff
date: 04/04/2022
output:
    html_notebook:
        code_folding: hide
---

# Loading packages

```{r load-packages}
############# PACKAGES AND OTHER PRELIMINARIES ############

# for easy loading of packages
if (!require("pacman")) install.packages("pacman")

# load packages
p_load("stringr", "ini", "gdata", "plyr", "dplyr", "here", "entropy")

# for relative file paths (e.g. here('Input', 'file.csv'))
here::i_am(file.path("scripts", "dataprep.rmd"))

# select correct here
here <- here::here

# Load custom helper functions
source(here("scripts", "helper_functions.R"))
```

# Loading data

```{r load-data}
####### EXPERIMENT CONFIG PARAMETERS #######

# Read in the experiment config file
config <- read.ini(here("data", "experiment", "raw", "config.txt"))
# Change data types
for (i in seq_along(config$experiment)) {
  config$experiment[i] <- as.double(config$experiment[i])
}
for (i in seq_along(config$training)) {
  config$training[i] <- as.double(config$training[i])
}

############# FILE LOCATIONS ############

# folders
data_participants_info_dir <- here("data", "experiment", "raw", "data_participants_info")
data_participants_tapping_dir <- here("data", "experiment", "raw", "data_participants_tapping")

# files
data_participants_info_filenames <- list.files(data_participants_info_dir)
data_participants_tapping_filenames <- list.files(data_participants_tapping_dir)

##### INITIAL DATAFRAME ######
pp_ids <- integer()

# First the general participant info file
if (length(data_participants_info_filenames) > 1) {
  stop("Error: More than one file in the data_participants_info dir")
} else {
  participants.info <-
    read.csv(file.path(
      data_participants_info_dir,
      data_participants_info_filenames[1]
    ))
}

# Then all the participant files
for (i in seq_along(data_participants_tapping_filenames)) {
  filename <- data_participants_tapping_filenames[i]
  if (endsWith(filename, ".csv")) {
    pp_id <- as.integer(str_extract(filename, "[0-9]+"))
    pp_ids <- c(pp_ids, pp_id)

    assign(
      sprintf("participant.tapping.%d", pp_id),
      read.csv(file.path(data_participants_tapping_dir, filename))
    )
  } else {
    stop("Files other than csv files in the participant tapping data dir")
  }
}

# Combine everything into one df

# start with an empty df
tapping.raw <- data.frame(matrix(nrow = 0, ncol = 30))
# get colnames from one of the df's and add column pp_id
colnames(tapping.raw) <- c("pp_id", colnames(get(paste0("participant.tapping.", pp_ids[1]))))

for (pp_id in pp_ids) {
  temp_list <- get(str_glue("participant.tapping.{pp_id}"))

  temp_df <- as.data.frame(temp_list)

  # add pp_id to df as column
  for (i in seq_len(nrow(temp_df))) {
    temp_df$pp_id[i] <- pp_id
  }
  # combine with empty list
  tapping.raw <- rbind(tapping.raw, temp_df)
}

# clean up
gdata::keep(list = c("calculatePDVs", "config", "participants.info", "pp_ids", "tapping.raw", "splitPythonListString"), sure = TRUE)
```

```{r prepare-participants-info}
# Could do some recoding here
```

# Cleaning

Remove all responses that have a different number of responses than the number of events in the stimulus:

```{r cleaning}
tapping.clean <- tapping.raw[0, ]
tapping.removed <- tapping.raw[0, ]

for (row_i in seq_len(nrow(tapping.raw))) {
  tapping_row <- tapping.raw[row_i, ]
  if (tapping_row$length == (length(splitPythonListString(tapping_row$resp_ITIs))) + 1) {
    tapping.clean <- rbind(tapping.clean, tapping_row)
  } else {
    tapping.removed <- rbind(tapping.removed, tapping_row)
  }
}
```


# ITI dataframe

```{r ITIs}
col_names <- c(
  "pp_id", "age", "gender", "L1", "handedness", "musical_experience", "stim_id", "sequence_id", "length", "n_resp", "stim_tempo_intended",
  "stim_tempo", "stim_sd", "interval_i", "stim_ioi", "resp_iti", "ioi_iti_diff"
)
ITIs <- data.frame(matrix(nrow = 0, ncol = length(col_names)))

for (row_i in seq_len(nrow(tapping.clean))) {
  tapping_row <- tapping.clean[row_i, ]
  pp_id <- tapping_row$pp_id
  print(pp_id)
  participant_row <- participants.info[participants.info$pp_id == pp_id, ]
  out_list <- as.list(participant_row)
  out_list$stim_id <- tapping_row$stim_id
  out_list$sequence_id <- paste(pp_id, tapping_row$stim_id, sep = "_")
  out_list$length <- tapping_row$length
  out_list$n_resp <- as.integer(length(splitPythonListString(tapping_row$resp_ITIs)) + 1)
  out_list$stim_tempo_intended <- tapping_row$int_tempo
  out_list$stim_tempo <- mean(diff(as.double(splitPythonListString(tapping_row$stim_onsets))))
  out_list$stim_sd <- sd(diff(as.double(splitPythonListString(tapping_row$stim_onsets))))
  out_list$interval_i <- 1:(out_list$length - 1)
  out_list$stim_ioi <- diff(as.double(splitPythonListString(tapping_row$stim_onsets)))
  out_list$resp_iti <- as.double(splitPythonListString(tapping_row$resp_ITIs))
  out_list$ioi_iti_diff <- abs(out_list$resp_iti - out_list$stim_ioi)

  ITIs <- rbind(ITIs, as.data.frame(out_list))
}

```


# Recoding

```{r recoding}

# Make everything capitalized
ITIs$gender <- str_to_title(ITIs$gender)
ITIs$L1 <- str_to_title(ITIs$L1)

# Recode gender
ITIs$gender <- dplyr::recode(ITIs$gender,
  "Woman" = "Female",
  "F" = "Female",
  "M" = "Male",
  "Man" = "Male",
  "Trans Woman" = "Trans woman",
  .missing = "ERROR")

# Recode L1
ITIs$L1 <- dplyr::recode(ITIs$L1,
  "German, Polish" = "German",
  "French " = "French"
)

```


# ITI by trial dataframe

```{r per-trial}
ITIs.bytrial <- ddply(ITIs, .(
  pp_id, age, gender, L1, handedness, musical_experience, stim_id, sequence_id,
  stim_tempo_intended, stim_tempo,
  stim_sd, length, n_resp
),
dplyr::summarize,
ITI_mean = mean(resp_iti),
ITI_sd = sd(resp_iti)
)

```




# save

```{r write-out}
write.csv(ITIs, here("data", "experiment", "processed", "ITIs.csv"), row.names = FALSE)
write.csv(ITIs.bytrial, here("data", "experiment", "processed", "ITIs_bytrial.csv"), row.names = FALSE)
write.csv(tapping.removed, here("data", "experiment", "processed", "removed.csv"), row.names = FALSE)
```

# Additional measures

Here we calculate additional measures using the scripts in the 'scripts/additional_measures' folder.
If not using Linux/Mac OS, you can run the scripts in that folder separately rather than using
the code below.

```{bash}

python3 ./scripts/additional_measures/calculate_editdistance.py
python3 ./scripts/additional_measures/calculate_entropy.py
python3 ./scripts/additional_measures/calculate_G_measure.py
python3 ./scripts/additional_measures/calculate_intervalratios.py
python3 ./scripts/additional_measures/calculate_rhythmic_contours.py
python3 ./scripts/additional_measures/ratio_preferences.py
python3 ./scripts/additional_measures/pp_measures.py


```