---
title: "NPOR: Plots"
author: jellevanderwerff
date: 04/04/2022
output:
    html_notebook:
        code_folding: hide
        toc: true
        toc_float: true
---

# Preliminaries

```{r load-packages, echo = FALSE, warning = FALSE, message = FALSE}

############# PACKAGES AND OTHER PRELIMINARIES ############

# for easy loading of packages
if (!require('pacman')) install.packages('pacman')

# load packages
p_load('stringr', 'ini', 'gdata', 'plyr', 'dplyr', 'here', 'ggplot2', 'reshape2', 'ggsci', 
'gghighlight', 'ggridges', 'viridis', 'RColorBrewer', 'lemon', 'scico')

# for relative file paths (e.g. here('Input', 'file.csv'))
here::i_am(file.path('scripts', 'plots', 'plots.Rmd'))

source(here('scripts', 'custom_ggplot_theme.R'))

# Knitr output settings
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  out.width = '100%'
)

labels <- as_labeller(c(
  '400' = '400 ms',
  '600' = '600 ms',
  '4' = '4 events',
  '5' = '5 events'
))

```


```{r load-data}

ITIs <- read.csv(here('data', 'experiment', 'processed', 'ITIs.csv'))
ITIs.bytrial <- read.csv(here('data', 'experiment', 'processed', 'ITIs_bytrial.csv'))
pp.measures <- read.csv(here('data', 'experiment', 'processed', 'by_pp.csv'))
interval.ratios <- read.csv(here('data', 'experiment', 'processed', 'interval_ratios.csv'))

```

```{r datatypes}

pp.measures$pp_id <- as.factor(pp.measures$pp_id)
pp.measures$stim_tempo_intended <- as.factor(pp.measures$stim_tempo_intended)
pp.measures$length <- as.factor(pp.measures$length)

pp.measures$condition <- as.factor(paste(pp.measures$stim_tempo_intended, pp.measures$length, sep = '_'))


ITIs$pp_id <- as.factor(ITIs$pp_id)
ITIs$condition <- as.factor(paste(pp.measures$stim_tempo_intended, pp.measures$length, sep = '_'))

ITIs.bytrial$condition <- as.factor(paste(ITIs.bytrial$stim_tempo_intended, ITIs.bytrial$length, sep = '_'))
ITIs.bytrial$pp_id <- as.factor(ITIs.bytrial$pp_id)


interval.ratios$pp_id <- as.factor(interval.ratios$pp_id)
interval.ratios$condition <- as.factor(paste(interval.ratios$stim_tempo_intended, interval.ratios$length, sep = '_'))

```


```




```

## Participant measures

###  G score

```{r g-score}

ggplot(pp.measures, aes(x = condition, y = G, colour = factor(pp_id), label = pp_id)) +
geom_point() +
geom_text(size = 8, vjust = 1, hjust = -0.3) +
theme_jelle()

```



### Tempo difference

```{r tempo-diff}

ggplot(pp.measures, aes(x = tempo_diff_avg, y = condition, colour = factor(pp_id), label = pp_id)) +
geom_pointrange(aes(xmin = tempo_diff_avg - tempo_diff_sd, xmax = tempo_diff_avg + tempo_diff_sd)) +
facet_grid(~condition)

```


### Entropy

```{r entropy}

ggplot(ITIs.bytrial, aes(y = entropy_diff, x = 1, colour = pp_id)) +
stat_summary(geom = "pointrange", position = position_dodge(width = 0.5)) +
facet_grid(rows = vars(stim_tempo_intended), cols = vars(length), labeller = labels) +
theme_jelle() +
theme(axis.title.x=element_blank(),
      axis.text.x=element_blank(),
      axis.ticks.x=element_blank()) +
labs(y = 'Entropy difference', title = "Entropy difference by participant and condition",
subtitle = "Negative values mean participants decreased entropy")

```')

```


## Distributions

### ITI ridges

```{r ITI-ridges}

ggplot(ITIs, aes(x = resp_iti, y = factor(pp_id), fill = factor(pp_id))) +
geom_density_ridges(alpha = 0.75) + 
theme_ridges() +
facet_grid(rows = vars(stim_tempo_intended), cols = vars(length), labeller = labels) +
theme_jelle() +
labs(x = 'ITI [ms]', y = 'Participant')

```

### Interval ratios

Here with the conditions plotted as facets:

```{r intervalratios-ridges}

ggplot(interval.ratios, aes(x = interval_ratio, y = factor(pp_id), fill = factor(pp_id))) +
geom_density_ridges(alpha = 0.75, show.legend = FALSE) + 
theme_ridges() +
facet_rep_grid(rows = vars(stim_tempo_intended), cols = vars(length), repeat.tick.labels = "x", 
        labeller = labels) +
theme_jelle() +
xlim(0, 1) +
scale_x_continuous(labels = c('', '1:3', '1:2', '1:1', '2:1', '3:1', ''), breaks = c(0, 0.25, 0.33, 0.5, 0.66, 0.75, 1)) +
geom_vline(xintercept = c(0.25, 0.33, 0.5, 0.66, 0.75), linetype = 'dashed', color = 'grey') +
labs(x = 'Interval ratio', y = 'Participant')

ggsave(here('plots', 'intervalratios-ridges.pdf'))

```


```{r intervalratios-ridges-condition}

ggplot(interval.ratios, aes(x = interval_ratio, y = factor(pp_id), fill = condition)) +
geom_density_ridges(alpha = 0.6) + 
theme_ridges() +
theme_jelle() +
xlim(0, 1) +
scale_x_continuous(labels = c('', '1:3', '1:2', '1:1', '2:1', '3:1', ''), breaks = c(0, 0.25, 0.33, 0.5, 0.66, 0.75, 1)) +
geom_vline(xintercept = c(0.25, 0.33, 0.5, 0.66, 0.75), linetype = 'dashed', color = 'grey') +
labs(x = 'Interval ratio', y = 'Participant') +
scale_fill_npg(labels = c('400 ms, 4 events', '400 ms, 5 events', '600 ms, 4 events', '600 ms, 5 events'))

ggsave(here('plots', 'intervalratios-ridges-2.pdf'))

```