---
title: "NPOR: Plots"
author: Jelle van der Werff
date: 08/08/2023
output:
    html_notebook:
        code_folding: hide
        toc: true
        toc_float: true
---


# Preliminaries

```{r load-packages, echo = FALSE, warning = FALSE, message = FALSE}
############# PACKAGES AND OTHER PRELIMINARIES ############

# for easy loading of packages
if (!require("pacman")) install.packages("pacman")

# load packages
p_load(
  "stringr", "ini", "gdata", "plyr", "dplyr", "here", "ggplot2", "reshape2", "ggsci",
  "gghighlight", "ggridges", "viridis", "RColorBrewer", "lemon", "scico", "latex2exp"
)

# for relative file paths (e.g. here('Input', 'file.csv'))
here::i_am(file.path("scripts", "plots", "plots.Rmd"))

source(here("scripts", "custom_ggplot_theme.R"))

# Knitr output settings
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  out.width = "100%",
  fig.width = 10,
  fig.height = 10,
  cache = TRUE
)

labels <- as_labeller(c(
  "400" = "400 ms",
  "600" = "600 ms",
  "4" = "4 events",
  "5" = "5 events"
))
```


```{r load-data}
ITIs <- read.csv(here("data", "experiment", "processed", "ITIs.csv"))
ITIs.bytrial <- read.csv(here("data", "experiment", "processed", "ITIs_bytrial.csv"))
pp.measures <- read.csv(here("data", "experiment", "processed", "pp_measures.csv"))
interval.ratios <- read.csv(here("data", "experiment", "processed", "interval_ratios.csv"))
interval.ratios.maxima <- read.csv(here("data", "experiment", "processed", "ratios_localmaxima.csv"))
```

```{r datatypes}
pp.measures$pp_id <- as.factor(pp.measures$pp_id)
pp.measures$stim_tempo_intended <- as.factor(pp.measures$stim_tempo_intended)
pp.measures$length <- as.factor(pp.measures$length)

pp.measures$condition <- as.factor(paste(pp.measures$stim_tempo_intended, pp.measures$length, sep = "_"))


ITIs$pp_id <- as.factor(ITIs$pp_id)
ITIs$condition <- as.factor(paste(ITIs$stim_tempo_intended, ITIs$length, sep = "_"))

ITIs.bytrial$condition <- as.factor(paste(ITIs.bytrial$stim_tempo_intended, ITIs.bytrial$length, sep = "_"))
ITIs.bytrial$pp_id <- as.factor(ITIs.bytrial$pp_id)


interval.ratios$pp_id <- as.factor(interval.ratios$pp_id)
interval.ratios$condition <- as.factor(paste(interval.ratios$stim_tempo_intended, interval.ratios$length, sep = "_"))

interval.ratios.maxima$pp_id <- as.factor(interval.ratios.maxima$pp_id)
interval.ratios.maxima$stim_tempo_intended <- as.factor(interval.ratios.maxima$stim_tempo_intended)
interval.ratios.maxima$length <- as.factor(interval.ratios.maxima$length)
interval.ratios.maxima$condition <- as.factor(paste(interval.ratios.maxima$stim_tempo_intended, interval.ratios.maxima$length, sep = "_"))
```


# Participant measures

##  G score

```{r g-score}
ggplot(pp.measures, aes(x = condition, y = G_resp, fill = condition, colour = condition)) +
  geom_violin(alpha = 0.7, show.legend = FALSE) +
  geom_point(size = 1, show.legend = FALSE, colour = "black", alpha = 0.3) +
  scale_x_discrete(labels = c(
    "400_4" = "400 ms\n4 events",
    "400_5" = "400 ms\n5 events",
    "600_4" = "600 ms\n4 events",
    "600_5" = "600 ms\n5 events"
  )) +
  stat_summary(geom = "pointrange", size = 1, linewidth = 1.5, colour = "black", show.legend = FALSE) +
  theme_jelle() +
  labs(
    x = "Condition",
    y = expression(italic(G) ~ measure),
    title = "Grammatical redundancy by condition",
    subtitle = "Higher values mean more grammatical redundancy",
    caption = "Interpretation: Let's say a participant always taps a quarternote and then an eighth note.
This means the quarternote-eighth note combination is redundant, as it could have been
one symbol, so G is higher. The more different combinations of symbols a participant uses,
the lower the G score."
  ) +
  theme(plot.caption = element_text(hjust = 0))



ggsave(here("plots", "g-score.pdf"))
```

## G bars
```{r g-bars}

ggplot(pp.measures, aes(x = pp_id, y = G_resp, fill = pp_id, colour = pp_id)) +
geom_bar(stat = "identity", position = position_dodge(width = 0.5), show.legend = FALSE) +
facet_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels, ) +
theme_jelle() +
labs(x = "Participant", y = "G measure (response)",
    title = "Grammatical redundancy by condition",
    subtitle = "Higher values mean more grammatical redundancy",
    caption = "Interpretation: Let's say a participant always taps a quarternote and then an eighth note.
This means the quarternote-eighth note combination is redundant, as it could have been
one symbol, so G is higher. The more different combinations of symbols a participant uses,
the lower the G score."
) +
theme(axis.text.x = element_blank())

ggsave(here("plots", "g-bars.pdf"))
```


## G difference

```{r g-diff-bycondition}
ggplot(pp.measures, aes(x = condition, y = G_diff, fill = condition, colour = condition)) +
  geom_violin(alpha = 0.7, show.legend = FALSE) +
  geom_point(size = 1, show.legend = FALSE, colour = "black", alpha = 0.3) +
  scale_x_discrete(labels = c(
    "400_4" = "400 ms\n4 events",
    "400_5" = "400 ms\n5 events",
    "600_4" = "600 ms\n4 events",
    "600_5" = "600 ms\n5 events"
  )) +
  stat_summary(geom = "pointrange", size = 1, linewidth = 1.5, colour = "black", show.legend = FALSE) +
  theme_jelle() +
  labs(
    x = "Condition",
    y = expression(italic(G)~"measure difference (response - stimulus)"),
    title = "Difference in grammatical redundancy by condition",
    subtitle = "Higher values mean more grammatical redundancy",
    caption = "Interpretation: Let's say a participant always taps a quarternote and then an eighth note.
This means the quarternote-eighth note combination is redundant, as it could have been
one symbol, so G is higher. The more different combinations of symbols a participant uses,
the lower the G score."
  ) +
  theme(plot.caption = element_text(hjust = 0))



ggsave(here("plots", "g-score-diff.pdf"))
```


## G difference

```{r g-diff}

# melt data
pp.measures.melt <- melt(pp.measures, id.vars = c("stim_tempo_intended", "length"), measure.vars = c("G_stim", "G_resp"))
levels(pp.measures.melt$variable) <- c("Stimulus", "Response")


ggplot(pp.measures.melt, aes(x = variable, y = value, colour = variable, group = variable)) +
geom_point(alpha = 0.2, show.legend = FALSE) +
stat_summary(geom = "pointrange", size = 1, linewidth = 1, show.legend = FALSE) +
facet_rep_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels, repeat.tick.labels = "x") +
theme_jelle() +
labs(
  y = expression(italic(G) ~ measure),
  title = "Grammatical redundancy by condition and stim/resp",
  subtitle = "Higher values mean more grammatical redundancy"
)

```

## Tempo difference

```{r tempo-diff}
pp.measures.std <- ddply(pp.measures, .(pp_id, stim_tempo_intended, length), summarize,
  y = tempo_diff_avg,
  ymin = tempo_diff_avg - tempo_diff_sd,
  ymax = tempo_diff_avg + tempo_diff_sd
)


ggplot(pp.measures.std, aes(x = 1, y = y, colour = factor(pp_id), label = pp_id)) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax), position = position_dodge(width = 0.1)) +
  theme_jelle() +
  facet_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  labs(
    y = "Tempo difference [ms]", title = "Tempo difference by participant and condition",
    subtitle = "Negative values mean participants tapped faster than the stimulus", colour = "Participant",
    caption = "Error bars indicate ± SD.\nNote that the values presented here are averages of
    positive and negative numbers, which complicates interpretation in terms of precision."
  ) +
  theme(legend.title = element_text(size = 12), legend.text = element_text(size = 12))

ggsave(here("plots", "tempo-diff.pdf"))
```


## Rhythmic contours edit distance

```{r rhythmic-contours}
ggplot(ITIs.bytrial, aes(x = rhythmic_contours_edit_distance, fill = condition)) +
  geom_bar(stat = "count", show.legend = FALSE) +
  facet_rep_grid(
    rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels,
    repeat.tick.labels = "x"
  ) +
  theme_jelle() +
  labs(
    x = "Number of edits", y = "Count", title = "Rhythmic contours edit distance by condition",
    caption = "The edit distance here is calculated by comparing two strings (one for the stimulus,
one for the response) of e.g. As and Bs, where the symbols indicate
whether an interval was longer or shorter than the previous one.",
    colour = "Participant"
  )
```

## Entropy

```{r entropy}
ggplot(ITIs.bytrial, aes(y = entropy_diff, x = 1, colour = pp_id)) +
  stat_summary(geom = "pointrange", position = position_dodge(width = 0.1)) +
  facet_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels) +
  theme_jelle() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  labs(
    y = "Entropy difference", title = "Entropy difference by participant and condition",
    subtitle = "Negative values mean participants decreased entropy", colour = "Participant"
  )
ggsave(here("plots", "entropy.pdf"))
```

## Preference for binary or ternary ratios

```{r binary-ternary-dataprep}
binary.ternary <- reshape2::melt(pp.measures,
  id.vars = c("pp_id", "stim_tempo_intended", "length"),
  measure.vars = c("preference_binary_D", "preference_ternary_D")
)

colnames(binary.ternary) <- c("pp_id", "stim_tempo_intended", "length", "ratio_type", "KS_D")
binary.ternary$ratio_type <- as.character(binary.ternary$ratio_type)
binary.ternary$ratio_type[binary.ternary$ratio_type == "preference_binary_D"] <- "Binary"
binary.ternary$ratio_type[binary.ternary$ratio_type == "preference_ternary_D"] <- "Ternary"
binary.ternary$ratio_type <- as.factor(binary.ternary$ratio_type)
```

```{r binary-ternary}
ggplot(binary.ternary, aes(x = ratio_type, y = KS_D, fill = ratio_type)) +
  geom_violin(alpha = 0.5, show.legend = FALSE) +
  geom_boxplot(width = 0.05, show.legend = FALSE, alpha = 0.2) +
  geom_point(size = 1, show.legend = FALSE, position = position_jitter(width = 0.2), color = "grey", alpha = 0.8) +
  facet_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels) +
  theme_jelle() +
  labs(
    x = "Ratio type", y = expression("Kolmigorov-Smirnov" ~ italic(D)), title = "Preference for binary vs. ternary ratios",
    subtitle = "Higher values mean more preference", colour = "Participant"
  )
ggsave(here("plots", "binary-ternary.pdf"))
```

# Distributions

## ITI ridges

```{r ITI-ridges, fig.width = 20, fig.height = 20}
ggplot(ITIs, aes(x = resp_iti, y = pp_id, fill = pp_id, colour = pp_id)) +
  geom_density_ridges(scale = 4, alpha = 0.9, show.legend = FALSE, color = "white") +
  facet_rep_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels, repeat.tick.labels = "x") +
  theme_jelle() +
  scale_x_continuous(breaks = seq(0, 1200, 200)) +
  coord_cartesian(xlim = c(0, 1200)) +
  labs(
    x = "ITI [ms]", y = "Participant", title = "Inter-tap intervals produced by participants by condition",
    subtitle = "Dashed line indicates (intended) tempo of condition"
  ) +
  geom_vline(data = ITIs[ITIs$stim_tempo_intended == 400, ], aes(xintercept = 400), linetype = "dashed", color = "grey") +
  geom_vline(data = ITIs[ITIs$stim_tempo_intended == 600, ], aes(xintercept = 600), linetype = "dashed", color = "grey")
ggsave(here("plots", "ITI-ridges.pdf"), width = 20, height = 20)
```

## Interval ratios

Here with the conditions plotted as facets:

```{r intervalratios-ridges, fig.width = 20, fig.height = 20}
ggplot(interval.ratios, aes(x = interval_ratio, y = factor(pp_id), fill = factor(pp_id))) +
  geom_density_ridges(scale = 4, alpha = 0.9, show.legend = FALSE, color = "white", from = 0.15, to = 0.85) +
  facet_rep_grid(
    rows = vars(length), cols = vars(stim_tempo_intended), repeat.tick.labels = "x",
    labeller = labels
  ) +
  theme_jelle() +
  coord_cartesian(xlim = c(0.15, 0.88)) +
  scale_x_continuous(labels = c("", "1:3", "1:2", "1:1", "2:1", "3:1", ""), breaks = c(0, 0.25, 0.33, 0.5, 0.66, 0.75, 1)) +
  geom_vline(xintercept = c(0.25, 0.33, 0.5, 0.66, 0.75), linetype = "dashed", color = "grey") +
  labs(
    x = "Interval ratio", y = "Participant",
    title = "Interval ratios by participant and condition",
    subtitle = TeX("Interval ratios calculated as $ITI_k = \\frac{ITI_k}{ITI_k + ITI_{k+1}}$"), fill = "Participant"
  )

ggsave(here("plots", "intervalratios-ridges.pdf"), width = 20, height = 20)
```


```{r quantized-intervalratios}

counts <- ddply(interval.ratios, .(condition, stim_tempo_intended, length, quantized_ratio_str), summarize, count = length(quantized_ratio_str))
counts <- counts[counts$count > 100, ]

counts <- counts[order(counts$condition, counts$count, decreasing = TRUE), ]
counts$quantized_ratio_str <- factor(counts$quantized_ratio_str, levels = unique(counts$quantized_ratio_str))


ggplot(counts, aes(x = quantized_ratio_str, y = count, fill = condition)) +
geom_bar(stat = "identity", show.legend = FALSE) +
facet_rep_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels, repeat.tick.labels = 'x') +
scale_fill_discrete(labels = c("400 ms; 4 events", "400 ms; 5 events", "600 ms; 4 events", "600 ms; 5 events")) +
theme_jelle() +
labs(x = "Interval ratio (quantized)", y = "Count", title = "Encountered interval ratios by condition",
subtitle = "Responses were first quantized to 1/16th notes.",
caption = "Only interval ratios encountered by more than 100 participants are shown.")

ggsave(here("plots", "quantized-intervalratios.pdf"))
```

```




## Local maxima

```{r ratios-localmaxima}
interval.ratios.summary <- ddply(interval.ratios.maxima, .(condition, stim_tempo_intended, length, n_maxima, maximum_i), summarize,
  maximum = mean(maximum_location),
  maximum_min_sd = mean(maximum_location) - sd(maximum_location),
  maximum_plus_sd = mean(maximum_location) + sd(maximum_location),
  n_observations = length(maximum_location)
)
interval.ratios.summary.n_observations <- ddply(interval.ratios.maxima, .(condition, stim_tempo_intended, length, n_maxima), dplyr::summarize,
  n = paste0(length(n_maxima), " pp's")
)


ggplot(interval.ratios.summary, aes(x = maximum, y = factor(n_maxima), xmin = maximum_min_sd, xmax = maximum_plus_sd, colour = condition)) +
  coord_cartesian(xlim = c(0, 1)) +
  geom_vline(xintercept = c(0.25, 0.33, 0.5, 0.66, 0.75), linetype = "dashed", color = "grey") +
  scale_x_continuous(labels = c("", "1:3", "1:2", "1:1", "2:1", "3:1", ""), breaks = c(0, 0.25, 0.33, 0.5, 0.66, 0.75, 1)) +
  geom_pointrange(show.legend = FALSE, size = 0.5, linewidth = 1) +
  facet_rep_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels, repeat.tick.labels = "x") +
  theme_jelle() +
  labs(
    x = "Location of local maximum (in integer ratios)",
    y = "Number of local maxima in participant's data",
    title = "Local maxima in interval ratio distributions",
    subtitle = "For a unimodal distribution, the number of local maxima is 1.\nFor a bimodal distribution, 2, etc.",
    caption = "Values averaged over participants"
  ) +
  geom_text(
    inherit.aes = FALSE, aes(x = 1, y = n_maxima, label = n),
    data = interval.ratios.summary.n_observations,
    nudge_y = 0.1,
    hjust = 1,
    size = 3,
    show.legend = FALSE,
    colour = "#cbcbcb"
  )

ggsave(here("plots", "ratios-localmaxima-summary.pdf"))
```


By pp:


```{r ratios-localmaxima-bypp, fig.height = 15}
# Get the number of local maxima for each participant in a dataframe with the pp_id
# and the number of local maxima
n_maxima <- ddply(interval.ratios.maxima, .(pp_id, condition, stim_tempo_intended, length), summarize, n_maxima = n_maxima[1])

ggplot(interval.ratios.maxima, aes(
  x = maximum_location, y = factor(pp_id), colour = condition
)) +
  geom_point(size = 2, show.legend = FALSE) +
  facet_rep_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels, repeat.tick.labels = "x") +
  scale_x_continuous(
    labels = c("", "1:3", "1:2", "1:1", "2:1", "3:1", ""),
    breaks = c(0, 0.25, 0.33, 0.5, 0.66, 0.75, 1)
  ) +
  theme_jelle() +
  geom_vline(xintercept = c(0.25, 0.33, 0.5, 0.66, 0.75), linetype = "dashed", color = "grey") +
  labs(
    x = "Location of local maximum", y = "Participant", title = "Local maxima in interval ratio distributions",
    subtitle = "Local maxima indicate the highest points in a participant's interval ratios distribution",
    caption = "Note that there may be different numbers of maxima for each participant/condition.
The number of maxima is given at the end of the participant's row."
  ) +
  geom_text(
    inherit.aes = FALSE,
    aes(x = 1, y = pp_id, label = n_maxima),
    data = n_maxima,
    size = 4,
    hjust = 0,
    vjust = 0.5,
    show.legend = FALSE
  ) +
  scale_colour_scico_d(palette = "roma", begin = 0)

ggsave(here("plots", "ratios-localmaxima-bypp.pdf"), height = 15, width = 10)
```