---
title: "NPOR: Plots"
author: Jelle van der Werff
date: 08/08/2023
output:
    html_notebook:
        code_folding: hide
        toc: true
        toc_float: true
---


# Preliminaries

```{r load-packages, echo = FALSE, warning = FALSE, message = FALSE}
############# PACKAGES AND OTHER PRELIMINARIES ############

# for easy loading of packages
if (!require("pacman")) install.packages("pacman")

# load packages
p_load(
  "stringr", "ini", "gdata", "plyr", "dplyr", "here", "ggplot2", "reshape2", "ggsci",
  "gghighlight", "ggridges", "viridis", "RColorBrewer", "lemon", "scico", "latex2exp", "GGally"
)

# for relative file paths (e.g. here('Input', 'file.csv'))
here::i_am(file.path("scripts", "plots", "plots.Rmd"))

source(here("scripts", "custom_ggplot_theme.R"))

# Knitr output settings
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  out.width = "100%",
  fig.width = 10,
  fig.height = 10,
  cache = TRUE
)

labels <- as_labeller(c(
  "400" = "400 ms",
  "600" = "600 ms",
  "4" = "4 events",
  "5" = "5 events"
))
```


```{r load-data}
ITIs <- read.csv(here("data", "experiment", "processed", "ITIs.csv"))
ITIs.bytrial <- read.csv(here("data", "experiment", "processed", "ITIs_bytrial.csv"))
pp.measures <- read.csv(here("data", "experiment", "processed", "pp_measures.csv"))
interval.ratios <- read.csv(here("data", "experiment", "processed", "interval_ratios.csv"))
interval.ratios.maxima <- read.csv(here("data", "experiment", "processed", "ratios_localmaxima.csv"))
```

```{r datatypes}
pp.measures$pp_id <- as.factor(pp.measures$pp_id)
pp.measures$stim_tempo_intended <- as.factor(pp.measures$stim_tempo_intended)
pp.measures$length <- as.factor(pp.measures$length)

pp.measures$condition <- as.factor(paste(pp.measures$stim_tempo_intended, pp.measures$length, sep = "_"))


ITIs$pp_id <- as.factor(ITIs$pp_id)
ITIs$condition <- as.factor(paste(ITIs$stim_tempo_intended, ITIs$length, sep = "_"))

ITIs.bytrial$condition <- as.factor(paste(ITIs.bytrial$stim_tempo_intended, ITIs.bytrial$length, sep = "_"))
ITIs.bytrial$pp_id <- as.factor(ITIs.bytrial$pp_id)


interval.ratios$pp_id <- as.factor(interval.ratios$pp_id)
interval.ratios$condition <- as.factor(paste(interval.ratios$stim_tempo_intended, interval.ratios$length, sep = "_"))

interval.ratios.maxima$pp_id <- as.factor(interval.ratios.maxima$pp_id)
interval.ratios.maxima$stim_tempo_intended <- as.factor(interval.ratios.maxima$stim_tempo_intended)
interval.ratios.maxima$length <- as.factor(interval.ratios.maxima$length)
interval.ratios.maxima$condition <- as.factor(paste(interval.ratios.maxima$stim_tempo_intended, interval.ratios.maxima$length, sep = "_"))
```


# Participant measures

##  G score

```{r g-score}
ggplot(pp.measures, aes(x = condition, y = G_resp, fill = condition, colour = condition)) +
  geom_violin(alpha = 0.7, show.legend = FALSE) +
  geom_point(size = 1, show.legend = FALSE, colour = "black", alpha = 0.3) +
  scale_x_discrete(labels = c(
    "400_4" = "400 ms\n4 events",
    "400_5" = "400 ms\n5 events",
    "600_4" = "600 ms\n4 events",
    "600_5" = "600 ms\n5 events"
  )) +
  stat_summary(geom = "pointrange", size = 1, linewidth = 1.5, colour = "black", show.legend = FALSE) +
  theme_jelle() +
  labs(
    x = "Condition",
    y = expression(italic(G) ~ measure),
    title = "Grammatical redundancy by condition",
    subtitle = "Higher values mean more grammatical redundancy",
    caption = "Interpretation: Let's say a participant always taps a quarternote and then an eighth note.
This means the quarternote-eighth note combination is redundant, as it could have been
one symbol, so G is higher. The more different combinations of symbols a participant uses,
the lower the G score."
  ) +
  theme(plot.caption = element_text(hjust = 0))



ggsave(here("plots", "g-score.pdf"))
```

## G bars
```{r g-bars}
ggplot(pp.measures, aes(x = pp_id, y = G_resp, fill = pp_id, colour = pp_id)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.5), show.legend = FALSE) +
  facet_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels, ) +
  theme_jelle() +
  labs(
    x = "Participant", y = "G measure (response)",
    title = "Grammatical redundancy by condition",
    subtitle = "Higher values mean more grammatical redundancy",
    caption = "Interpretation: Let's say a participant always taps a quarternote and then an eighth note.
This means the quarternote-eighth note combination is redundant, as it could have been
one symbol, so G is higher. The more different combinations of symbols a participant uses,
the lower the G score."
  ) +
  theme(axis.text.x = element_blank())

ggsave(here("plots", "g-bars.pdf"))
```


## G difference

```{r g-diff-bycondition}
ggplot(pp.measures, aes(x = condition, y = G_diff, fill = condition, colour = condition)) +
  geom_violin(alpha = 0.7, show.legend = FALSE) +
  geom_point(size = 1, show.legend = FALSE, colour = "black", alpha = 0.3) +
  scale_x_discrete(labels = c(
    "400_4" = "400 ms\n4 events",
    "400_5" = "400 ms\n5 events",
    "600_4" = "600 ms\n4 events",
    "600_5" = "600 ms\n5 events"
  )) +
  stat_summary(geom = "pointrange", size = 1, linewidth = 1.5, colour = "black", show.legend = FALSE) +
  theme_jelle() +
  labs(
    x = "Condition",
    y = expression(italic(G) ~ "measure difference (response - stimulus)"),
    title = "Difference in grammatical redundancy by condition",
    subtitle = "Higher values mean more grammatical redundancy",
    caption = "Interpretation: Let's say a participant always taps a quarternote and then an eighth note.
This means the quarternote-eighth note combination is redundant, as it could have been
one symbol, so G is higher. The more different combinations of symbols a participant uses,
the lower the G score."
  ) +
  theme(plot.caption = element_text(hjust = 0))



ggsave(here("plots", "g-score-diff.pdf"))
```


## G difference

```{r g-diff}
# melt data
pp.measures.melt <- melt(pp.measures, id.vars = c("stim_tempo_intended", "length"), measure.vars = c("G_stim", "G_resp"))
levels(pp.measures.melt$variable) <- c("Stimulus", "Response")


ggplot(pp.measures.melt, aes(x = variable, y = value, colour = variable, group = variable)) +
  geom_point(alpha = 0.2, show.legend = FALSE) +
  stat_summary(geom = "pointrange", size = 1, linewidth = 1, show.legend = FALSE) +
  facet_rep_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels, repeat.tick.labels = "x") +
  theme_jelle() +
  labs(
    y = expression(italic(G) ~ measure),
    title = "Grammatical redundancy by condition and stim/resp",
    subtitle = "Higher values mean more grammatical redundancy"
  )
```

## Tempo difference

```{r tempo-diff}
pp.measures.std <- ddply(pp.measures, .(pp_id, stim_tempo_intended, length), summarize,
  y = tempo_diff_avg,
  ymin = tempo_diff_avg - tempo_diff_sd,
  ymax = tempo_diff_avg + tempo_diff_sd
)


ggplot(pp.measures.std, aes(x = 1, y = y, colour = factor(pp_id), label = pp_id)) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax), position = position_dodge(width = 0.1)) +
  theme_jelle() +
  facet_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  labs(
    y = "Tempo difference [ms]", title = "Tempo difference by participant and condition",
    subtitle = "Negative values mean participants tapped faster than the stimulus", colour = "Participant",
    caption = "Error bars indicate ± SD.\nNote that the values presented here are averages of
    positive and negative numbers, which complicates interpretation in terms of precision."
  ) +
  theme(legend.title = element_text(size = 12), legend.text = element_text(size = 12))

ggsave(here("plots", "tempo-diff.pdf"))
```


## Entropy

```{r entropy}
ggplot(ITIs.bytrial, aes(y = entropy_diff, x = 1, colour = pp_id)) +
  stat_summary(geom = "pointrange", position = position_dodge(width = 0.1)) +
  facet_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels) +
  theme_jelle() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  labs(
    y = "Entropy difference", title = "Entropy difference by participant and condition",
    subtitle = "Negative values mean participants decreased entropy", colour = "Participant"
  )
ggsave(here("plots", "entropy.pdf"))
```

## Preference for small integer ratios

```{r simple-integers}
ggplot(pp.measures, aes(x = condition, y = small_integers_vs_total_prop, colour = condition, group = condition)) +
  geom_point(show.legend = FALSE, alpha = 0.3) +
  stat_summary(geom = "pointrange", show.legend = FALSE, size = 1, linewidth = 2) +
  scale_x_discrete(labels = c(
    "400_4" = "400 ms\n4 events",
    "400_5" = "400 ms\n5 events",
    "600_4" = "600 ms\n4 events",
    "600_5" = "600 ms\n5 events"
  )) +
  labs(
    x = "Condition", y = "Proportion of simple integer ratios",
    title = "Proportion of simple integer ratios by condition",
    subtitle = "Each point represents one participant",
    caption = "Simple integer ratios were all combinations of the numbers 1 to 5. For example, 1:1, 3:5, 5:3, 2:1, etc."
  ) +
  theme_jelle()

ggsave(here("plots", "simple_integers.pdf"))
ggsave(here("plots", "simple_integers.png"), dpi = 300)
```


## Edit distance

### Across participants

```{r edit-distance}
ggplot(ITIs.bytrial, aes(x = edit_distance_normalized, colour = condition, fill = condition, group = condition)) +
  geom_histogram(show.legend = FALSE, alpha = 0.8) +
  labs(
    x = "Condition", y = "Count",
    title = "Normalized edit distance by condition across participants",
    caption = "Edit distance is the number of insertions, deletions, and substitutions needed to transform one sequence into another."
  ) +
  facet_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels) +
  theme_jelle()

ggsave(here("plots", "edit_distance.pdf"))
ggsave(here("plots", "edit_distance.png"), dpi = 300)
```

### By participant

```{r edit-distance-bypp}
ggplot(pp.measures, aes(x = condition, y = edit_distance_norm_avg, colour = condition, fill = condition, group = condition)) +
  geom_point(show.legend = FALSE, alpha = 0.8) +
  stat_summary(geom = "pointrange", show.legend = FALSE, size = 1, linewidth = 2) +
  labs(
    x = "Condition", y = "Edit distance (normalized)",
    title = "Normalized edit distance by participant and condition",
    subtitle = "Each point represents one participant",
    caption = "Edit distance is the number of insertions, deletions, and substitutions needed to transform one sequence into another.
    Normalization was performed by dividing by the maximum possible edit distance."
  ) +
  scale_x_discrete(labels = c(
    "400_4" = "400 ms\n4 events",
    "400_5" = "400 ms\n5 events",
    "600_4" = "600 ms\n4 events",
    "600_5" = "600 ms\n5 events"
  )) +
  theme_jelle()

ggsave(here("plots", "edit_distance_bypp.pdf"))
ggsave(here("plots", "edit_distance_bypp.png"), dpi = 300)
```

# Distributions

## ITI ridges

```{r ITI-ridges, fig.width = 20, fig.height = 20}
ggplot(ITIs, aes(x = resp_iti, y = pp_id, fill = pp_id, colour = pp_id)) +
  geom_density_ridges(scale = 4, alpha = 0.9, show.legend = FALSE, color = "white") +
  facet_rep_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels, repeat.tick.labels = "x") +
  theme_jelle() +
  scale_x_continuous(breaks = seq(0, 1200, 200)) +
  coord_cartesian(xlim = c(0, 1200)) +
  labs(
    x = "ITI [ms]", y = "Participant", title = "Inter-tap intervals produced by participants by condition",
    subtitle = "Dashed line indicates (intended) tempo of condition"
  ) +
  geom_vline(data = ITIs[ITIs$stim_tempo_intended == 400, ], aes(xintercept = 400), linetype = "dashed", color = "grey") +
  geom_vline(data = ITIs[ITIs$stim_tempo_intended == 600, ], aes(xintercept = 600), linetype = "dashed", color = "grey")
ggsave(here("plots", "ITI-ridges.pdf"), width = 20, height = 20)
```

## Interval ratios

Here with the conditions plotted as facets:

```{r intervalratios-ridges, fig.width = 20, fig.height = 20}
ggplot(interval.ratios, aes(x = interval_ratio, y = factor(pp_id), fill = factor(pp_id))) +
  geom_density_ridges(scale = 4, alpha = 0.9, show.legend = FALSE, color = "white", from = 0.15, to = 0.85) +
  facet_rep_grid(
    rows = vars(length), cols = vars(stim_tempo_intended), repeat.tick.labels = "x",
    labeller = labels
  ) +
  theme_jelle() +
  coord_cartesian(xlim = c(0.15, 0.88)) +
  scale_x_continuous(labels = c("", "1:3", "1:2", "1:1", "2:1", "3:1", ""), breaks = c(0, 0.25, 0.33, 0.5, 0.66, 0.75, 1)) +
  geom_vline(xintercept = c(0.25, 0.33, 0.5, 0.66, 0.75), linetype = "dashed", color = "grey") +
  labs(
    x = "Interval ratio", y = "Participant",
    title = "Interval ratios by participant and condition",
    subtitle = TeX("Interval ratios calculated as $ITI_k = \\frac{ITI_k}{ITI_k + ITI_{k+1}}$"), fill = "Participant"
  )

ggsave(here("plots", "intervalratios-ridges.pdf"), width = 20, height = 20)
```


```{r quantized-intervalratios}
counts <- ddply(interval.ratios, .(condition, stim_tempo_intended, length, quantized_ratio_str), summarize, count = length(quantized_ratio_str))
counts <- counts[counts$count > 100, ]

counts <- counts[order(counts$condition, counts$count, decreasing = TRUE), ]
counts$quantized_ratio_str <- factor(counts$quantized_ratio_str, levels = unique(counts$quantized_ratio_str))


ggplot(counts, aes(x = quantized_ratio_str, y = count, fill = condition)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_rep_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels, repeat.tick.labels = "x") +
  scale_fill_discrete(labels = c("400 ms; 4 events", "400 ms; 5 events", "600 ms; 4 events", "600 ms; 5 events")) +
  theme_jelle() +
  labs(
    x = "Interval ratio (quantized)", y = "Count", title = "Encountered interval ratios by condition",
    subtitle = "Responses were first quantized to 1/16th notes.",
    caption = "Only interval ratios encountered by more than 100 participants are shown."
  )

ggsave(here("plots", "quantized-intervalratios.pdf"))
```






## Local maxima

```{r ratios-localmaxima}
interval.ratios.summary <- ddply(interval.ratios.maxima, .(condition, stim_tempo_intended, length, n_maxima, maximum_i), summarize,
  maximum = mean(maximum_location),
  maximum_min_sd = mean(maximum_location) - sd(maximum_location),
  maximum_plus_sd = mean(maximum_location) + sd(maximum_location),
  n_observations = length(maximum_location)
)
interval.ratios.summary.n_observations <- ddply(interval.ratios.maxima, .(condition, stim_tempo_intended, length, n_maxima), dplyr::summarize,
  n = paste0(length(n_maxima), " pp's")
)


ggplot(interval.ratios.summary, aes(x = maximum, y = factor(n_maxima), xmin = maximum_min_sd, xmax = maximum_plus_sd, colour = condition)) +
  coord_cartesian(xlim = c(0, 1)) +
  geom_vline(xintercept = c(0.25, 0.33, 0.5, 0.66, 0.75), linetype = "dashed", color = "grey") +
  scale_x_continuous(labels = c("", "1:3", "1:2", "1:1", "2:1", "3:1", ""), breaks = c(0, 0.25, 0.33, 0.5, 0.66, 0.75, 1)) +
  geom_pointrange(show.legend = FALSE, size = 0.5, linewidth = 1) +
  facet_rep_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels, repeat.tick.labels = "x") +
  theme_jelle() +
  labs(
    x = "Location of local maximum (in integer ratios)",
    y = "Number of local maxima in participant's data",
    title = "Local maxima in interval ratio distributions",
    subtitle = "For a unimodal distribution, the number of local maxima is 1.\nFor a bimodal distribution, 2, etc.",
    caption = "Values averaged over participants"
  ) +
  geom_text(
    inherit.aes = FALSE, aes(x = 1, y = n_maxima, label = n),
    data = interval.ratios.summary.n_observations,
    nudge_y = 0.1,
    hjust = 1,
    size = 3,
    show.legend = FALSE,
    colour = "#cbcbcb"
  )

ggsave(here("plots", "ratios-localmaxima-summary.pdf"))
```


By pp:


```{r ratios-localmaxima-bypp, fig.height = 15}
# Get the number of local maxima for each participant in a dataframe with the pp_id
# and the number of local maxima
n_maxima <- ddply(interval.ratios.maxima, .(pp_id, condition, stim_tempo_intended, length), summarize, n_maxima = n_maxima[1])

ggplot(interval.ratios.maxima, aes(
  x = maximum_location, y = factor(pp_id), colour = condition
)) +
  geom_point(size = 2, show.legend = FALSE) +
  facet_rep_grid(rows = vars(length), cols = vars(stim_tempo_intended), labeller = labels, repeat.tick.labels = "x") +
  scale_x_continuous(
    labels = c("", "1:3", "1:2", "1:1", "2:1", "3:1", ""),
    breaks = c(0, 0.25, 0.33, 0.5, 0.66, 0.75, 1)
  ) +
  theme_jelle() +
  geom_vline(xintercept = c(0.25, 0.33, 0.5, 0.66, 0.75), linetype = "dashed", color = "grey") +
  labs(
    x = "Location of local maximum", y = "Participant", title = "Local maxima in interval ratio distributions",
    subtitle = "Local maxima indicate the highest points in a participant's interval ratios distribution",
    caption = "Note that there may be different numbers of maxima for each participant/condition.
The number of maxima is given at the end of the participant's row."
  ) +
  geom_text(
    inherit.aes = FALSE,
    aes(x = 1, y = pp_id, label = n_maxima),
    data = n_maxima,
    size = 4,
    hjust = 0,
    vjust = 0.5,
    show.legend = FALSE
  ) +
  scale_colour_scico_d(palette = "roma", begin = 0)

ggsave(here("plots", "ratios-localmaxima-bypp.pdf"), height = 15, width = 10)
```



```{r one-big-plot}
edit_distance <- ggplot(pp.measures, aes(x = condition, y = edit_distance_norm_avg, colour = condition, fill = condition, group = condition)) +
  geom_point(show.legend = FALSE, alpha = 0.8) +
  stat_summary(geom = "pointrange", show.legend = FALSE, size = 1, linewidth = 2) +
  labs(
    y = "Edit distance (normalized)",
  ) +
  scale_x_discrete(labels = c(
    "400_4" = "400 ms\n4 events",
    "400_5" = "400 ms\n5 events",
    "600_4" = "600 ms\n4 events",
    "600_5" = "600 ms\n5 events"
  )) +
  theme_jelle() +
  theme(axis.title.x = element_blank())

G_resp <- ggplot(pp.measures, aes(x = condition, y = G_resp, colour = condition, fill = condition, group = condition)) +
  geom_point(show.legend = FALSE, alpha = 0.8) +
  stat_summary(geom = "pointrange", show.legend = FALSE, size = 1, linewidth = 2) +
  labs(
    y = "G (response)"
  ) +
  scale_x_discrete(labels = c(
    "400_4" = "400 ms\n4 events",
    "400_5" = "400 ms\n5 events",
    "600_4" = "600 ms\n4 events",
    "600_5" = "600 ms\n5 events"
  )) +
  theme_jelle() +
  theme(axis.title.x = element_blank())


entropy_diff <- ggplot(pp.measures, aes(x = condition, y = entropy_diff_avg, colour = condition, fill = condition, group = condition)) +
  geom_point(show.legend = FALSE, alpha = 0.8) +
  stat_summary(geom = "pointrange", show.legend = FALSE, size = 1, linewidth = 2) +
  labs(
    y = "Entropy difference"
  ) +
  scale_x_discrete(labels = c(
    "400_4" = "400 ms\n4 events",
    "400_5" = "400 ms\n5 events",
    "600_4" = "600 ms\n4 events",
    "600_5" = "600 ms\n5 events"
  )) +
  theme(axis.title.x = element_blank()) +
  theme_jelle()

integer_ratios <- ggplot(pp.measures, aes(x = condition, y = small_integers_vs_total_prop, colour = condition, fill = condition, group = condition)) +
  geom_point(show.legend = FALSE, alpha = 0.8) +
  stat_summary(geom = "pointrange", show.legend = FALSE, size = 1, linewidth = 2) +
  labs(
    y = "Small integer ratios proportion"
  ) +
  scale_x_discrete(labels = c(
    "400_4" = "400 ms\n4 events",
    "400_5" = "400 ms\n5 events",
    "600_4" = "600 ms\n4 events",
    "600_5" = "600 ms\n5 events"
  )) +
  theme(axis.title.x = element_blank()) +
  theme_jelle()

# combined cowplot
plot_row <- cowplot::plot_grid(
  edit_distance, G_resp, entropy_diff, integer_ratios,
  ncol = 2,
  nrow = 2
  # axis = "tblr"
  # rel_widths = c(1, 1),
  # rel_heights = c(1, 1)
)

ggsave(here("plots", "one-big-plot.pdf"), width = 10, height = 10, scale = 1.5)
ggsave(here("plots", "one-big-plot.png"), width = 10, height = 10, scale = 1.5)
```

```{r one-big-plot-2}
# Melt dataframe
pp.measures.melt <- melt(pp.measures, id.vars = "condition", measure.vars = c(
  "G_resp", "edit_distance_norm_avg", "entropy_diff_avg",
  "small_integers_vs_total_prop"
))
levels(pp.measures.melt$variable) <- c("G (response)", "Edit distance", "Entropy difference", "Small integer ratios proportion")
levels(pp.measures.melt$condition) <- c("400 ms\n4 events", "400 ms\n5 events", "600 ms\n4 events", "600 ms\n5 events")
colnames(pp.measures.melt) <- c("Condition", "Measure", "Value")

# Plot
ggplot(pp.measures.melt, aes(x = Condition, y = Value, colour = Condition, fill = Condition, group = Condition)) +
  geom_point(show.legend = FALSE, alpha = 0.8) +
  geom_violin(show.legend = FALSE, alpha = 0.5) +
  stat_summary(geom = "pointrange", show.legend = FALSE, size = 1, linewidth = 2) +
  facet_rep_wrap(~Measure, ncol = 2, nrow = 2, scales = "free_y", repeat.tick.labels = "x") +
  labs(title = "Participant measures by condition", subtitle = "Each point represents one participant") +
  theme(axis.title.y = element_blank(), plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
  theme_jelle()

ggsave(here("plots", "one-big-plot-2.pdf"), width = 10, height = 10, scale = 1.5)
ggsave(here("plots", "one-big-plot-2.png"), width = 10, height = 10, scale = 1.5)
```


# Correlations

```{r correlations}
labels <- as_labeller(c(
  "G_resp" = "G (response)",
  "edit_distance_norm_avg" = "Edit distance",
  "entropy_diff_avg" = "Entropy difference",
  "rhythmic_contours_edit_distance_quantized_avg" = "Contours edit distance (qntzd)",
  "small_integers_vs_total_prop" = "Small integer ratios proportion"
))

levels(pp.measures$condition) <- c("400 ms (4)", "400 ms (5)", "600 ms (4)", "600 ms (5)")

# Scatterplots with combinations between all measures
ggplot2::theme_set(theme_minimal())
ggpairs(pp.measures,
  columns = c("G_resp", "edit_distance_norm_avg", "entropy_diff_avg", "rhythmic_contours_edit_distance_quantized_avg", "small_integers_vs_total_prop"),
  title = "Correlations between participant measures", labeller = labels, aes(colour = condition, alpha = 0.8)
)

ggsave(here("plots", "correlations", "correlations.pdf"))
ggsave(here("plots", "correlations", "correlations.png"), bg = "white")
```



small integer ratios and entropy difference:

```{r integerratios-vs-entropydiff}
labels <- as_labeller(c(
  "400_4" = "400 ms\n4 events",
  "400_5" = "400 ms\n5 events",
  "600_4" = "600 ms\n4 events",
  "600_5" = "600 ms\n5 events"
))

ggplot(pp.measures, aes(x = entropy_diff_avg, y = small_integers_vs_total_prop, colour = condition, fill = condition, group = condition)) +
  geom_point(show.legend = FALSE, alpha = 0.8) +
  geom_smooth(show.legend = FALSE, method = "lm") +
  facet_rep_wrap(~condition, ncol = 2, nrow = 2, scales = "free_y", repeat.tick.labels = "x") +
  labs(x = "Entropy difference", y = "Small integer ratios proportion") +
  theme_jelle()

ggsave(here("plots", "correlations", "integerratios-vs-entropydiff.pdf"))
ggsave(here("plots", "correlations", "integerratios-vs-entropydiff.png"))
```

```{r g-to-integerratios}
labels <- as_labeller(c(
  "400_4" = "400 ms\n4 events",
  "400_5" = "400 ms\n5 events",
  "600_4" = "600 ms\n4 events",
  "600_5" = "600 ms\n5 events"
))

ggplot(pp.measures, aes(x = G_resp, y = small_integers_vs_total_prop, colour = condition, fill = condition, group = condition)) +
  geom_point(show.legend = FALSE, alpha = 0.8) +
  geom_smooth(show.legend = FALSE, method = "lm") +
  facet_rep_wrap(~condition, ncol = 2, nrow = 2, scales = "free", repeat.tick.labels = "x") +
  labs(x = "G", y = "Small integer ratios proportion") +
  theme_jelle()

ggsave(here("plots", "correlations", "G-vs-integerratios.pdf"))
ggsave(here("plots", "correlations", "G-vs-integerratios.png"))
```


```{r entropy-to-editdistance}
labels <- as_labeller(c(
  "400_4" = "400 ms\n4 events",
  "400_5" = "400 ms\n5 events",
  "600_4" = "600 ms\n4 events",
  "600_5" = "600 ms\n5 events"
))

ggplot(pp.measures, aes(x = abs(entropy_diff_avg), y = edit_distance_norm_avg, colour = condition, fill = condition, group = condition)) +
  geom_point(show.legend = FALSE, alpha = 0.8) +
  geom_smooth(show.legend = FALSE, method = "lm") +
  facet_rep_wrap(~condition, ncol = 2, nrow = 2, scales = "free_y", repeat.tick.labels = "x") +
  labs(x = "Absolute entropy difference", y = "Edit distance") +
  theme_jelle()

ggsave(here("plots", "correlations", "entropy-vs-editdistance-abs.pdf"))
ggsave(here("plots", "correlations", "entropy-vs-editdistance-abs.png"))
```

```{r entropy-to-editdistance}
labels <- as_labeller(c(
  "400_4" = "400 ms\n4 events",
  "400_5" = "400 ms\n5 events",
  "600_4" = "600 ms\n4 events",
  "600_5" = "600 ms\n5 events"
))

ggplot(pp.measures, aes(x = entropy_resp_avg, y = edit_distance_norm_avg, colour = condition, fill = condition, group = condition)) +
  geom_point(show.legend = FALSE, alpha = 0.8) +
  facet_rep_wrap(~condition, ncol = 2, nrow = 2, repeat.tick.labels = "x") +
  labs(x = "Entropy (response)", y = "Edit distance") +
  theme_jelle()

ggsave(here("plots", "correlations", "response_entropy_vs_editdistance.pdf"))
ggsave(here("plots", "correlations", "response_entropy_vs_editdistance.png"))
```


# 2D plots

```{r editdistance-vs-smallintegers-2d}
ggplot(pp.measures, aes(x = edit_distance_norm_avg, y = small_integers_vs_total_prop)) +
  geom_density_2d_filled(alpha = 0.8, show.legend = FALSE) +
  theme_jelle() +
  labs(
    x = "Edit distance (normalized)",
    y = "Small integer ratios proportion of total",
    title = "Edit distance vs proportion of small integer ratios",
    subtitle = "Higher values for edit distance mean more difference between stimulus and response
Higher values for small integer ratios mean more small integer ratios in the response",
    caption = "Edit distance is the number of insertions, deletions, and substitutions needed to transform one sequence into another.
Normalization was performed by dividing by the maximum possible edit distance."
  ) +
  theme(plot.caption = element_text(hjust = 0))


ggsave(here("plots", "correlations", "editdistance-vs-smallintegers-2d.pdf"))
ggsave(here("plots", "correlations", "editdistance-vs-smallintegers-2d.png"))
```


```{r editdistance-vs-G-2d}
ggplot(pp.measures, aes(x = edit_distance_norm_avg, y = G_resp)) +
  geom_density_2d_filled(alpha = 0.8, show.legend = FALSE) +
  theme_jelle() +
  labs(
    x = "Edit distance", y = "G", title = "Edit distance vs grammatical redundancy",
    subtitle = "Higher values for edit distance mean more difference between stimulus and response
Higher values for G mean more grammatical structure in the response"
  )

ggsave(here("plots", "correlations", "editdistance-vs-G-2d.pdf"))
ggsave(here("plots", "correlations", "editdistance-vs-G-2d.png"))
```

```{r editdistance-vs-entropydiff-2d}

ggplot(pp.measures, aes(x = edit_distance_norm_avg, y = entropy_diff_avg)) +
  geom_density_2d_filled(alpha = 0.8, show.legend = FALSE) +
  theme_jelle() +
  labs(
    x = "Edit distance", y = "Entropy difference", title = "Edit distance vs entropy difference",
    subtitle = "Higher values for edit distance mean more difference between stimulus and response
Negative values for entropy difference mean more decrease in entropy in the response")

ggsave(here("plots", "correlations", "editdistance-vs-entropydiff-2d.pdf"))
ggsave(here("plots", "correlations", "editdistance-vs-entropydiff-2d.png"))
```




```{r entropydiff-vs-G-2d}

ggplot(pp.measures, aes(x = entropy_diff_avg, y = G_resp)) +
  geom_density_2d_filled(alpha = 0.8, show.legend = FALSE) +
  theme_jelle() +
  labs(
    x = "Entropy difference", y = "G", title = "Entropy difference vs grammatical redundancy",
    subtitle = "Negative values for entropy difference mean more decrease in entropy in the response
Higher values for G mean more grammatical redundancy in the response"
  )

ggsave(here("plots", "correlations", "entropydiff-vs-G-2d.pdf"))
ggsave(here("plots", "correlations", "entropydiff-vs-G-2d.png"))

```

```{r entropydiff-vs-smallintegers-2d}

ggplot(pp.measures, aes(x = entropy_diff_avg, y = small_integers_vs_total_prop)) +
  geom_density_2d_filled(alpha = 0.8, show.legend = FALSE) +
  theme_jelle() +
  labs(
    x = "Entropy difference", y = "Small integer ratios proportion of total",
    title = "Entropy difference vs proportion of small integer ratios",
    subtitle = "Negative values for entropy difference mean more decrease in entropy in the response
Higher values for small integer ratios mean more small integer ratios in the response"
  )

ggsave(here("plots", "correlations", "entropydiff-vs-smallintegers-2d.pdf"))
ggsave(here("plots", "correlations", "entropydiff-vs-smallintegers-2d.png"))

```

```


```{r G-vs-smallintegers-2d}

ggplot(pp.measures, aes(x = G_resp, y = small_integers_vs_total_prop)) +
  geom_density_2d_filled(alpha = 0.8, show.legend = FALSE) +
  theme_jelle() +
  labs(
    x = "G", y = "Small integer ratios proportion of total",
    title = "Grammatical redundancy vs proportion of small integer ratios",
    subtitle = "Higher values for G mean more grammatical redundancy in the response
Higher values for small integer ratios mean more small integer ratios in the response"
  )

ggsave(here("plots", "correlations", "G-vs-smallintegers-2d.pdf"))
ggsave(here("plots", "correlations", "G-vs-smallintegers-2d.png"))

```












Just as an illustration, let's see what two variables look like in such a plot that share a linear relationship:

```{r linear-relationship}

# Plot entropy difference against entropy response

ggplot(pp.measures, aes(x = entropy_resp_avg, y = entropy_diff_avg)) +
  geom_density_2d_filled(alpha = 0.8, show.legend = FALSE) +
  theme_jelle()
```


