---
title: "NPOR: Plots"
author: jellevanderwerff
date: 04/04/2022
output:
    html_notebook:
        code_folding: hide
        toc: true
        toc_float: true
---

# Preliminaries

```{r load-packages, echo = FALSE, warning = FALSE, message = FALSE}
############# PACKAGES AND OTHER PRELIMINARIES ############

# for easy loading of packages
if (!require("pacman")) install.packages("pacman")

# load packages
p_load(
  "stringr", "ini", "gdata", "plyr", "dplyr", "here", "ggplot2", "reshape2", "ggsci",
  "gghighlight", "ggridges", "viridis", "RColorBrewer", "lemon", "scico"
)

# for relative file paths (e.g. here('Input', 'file.csv'))
here::i_am(file.path("scripts", "plots", "plots.Rmd"))

source(here("scripts", "custom_ggplot_theme.R"))

# Knitr output settings
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  out.width = "100%",
  cache = TRUE
)

labels <- as_labeller(c(
  "400" = "400 ms",
  "600" = "600 ms",
  "4" = "4 events",
  "5" = "5 events"
))
```


```{r load-data}
ITIs <- read.csv(here("data", "experiment", "processed", "ITIs.csv"))
ITIs.bytrial <- read.csv(here("data", "experiment", "processed", "ITIs_bytrial.csv"))
pp.measures <- read.csv(here("data", "experiment", "processed", "by_pp.csv"))
interval.ratios <- read.csv(here("data", "experiment", "processed", "interval_ratios.csv"))
```

```{r datatypes}
pp.measures$pp_id <- as.factor(pp.measures$pp_id)
pp.measures$stim_tempo_intended <- as.factor(pp.measures$stim_tempo_intended)
pp.measures$length <- as.factor(pp.measures$length)

pp.measures$condition <- as.factor(paste(pp.measures$stim_tempo_intended, pp.measures$length, sep = "_"))


ITIs$pp_id <- as.factor(ITIs$pp_id)
ITIs$condition <- as.factor(paste(pp.measures$stim_tempo_intended, pp.measures$length, sep = "_"))

ITIs.bytrial$condition <- as.factor(paste(ITIs.bytrial$stim_tempo_intended, ITIs.bytrial$length, sep = "_"))
ITIs.bytrial$pp_id <- as.factor(ITIs.bytrial$pp_id)


interval.ratios$pp_id <- as.factor(interval.ratios$pp_id)
interval.ratios$condition <- as.factor(paste(interval.ratios$stim_tempo_intended, interval.ratios$length, sep = "_"))
```


## Participant measures

###  G score

```{r g-score}
ggplot(pp.measures, aes(x = condition, y = G, fill = condition)) +
  geom_point(size = 1, show.legend = FALSE) +
  scale_x_discrete(labels = c("400_4" = "400 ms\n4 events", "400_5" = "400 ms\n5 events", "600_4" = "600 ms\n4 events", "600_5" = "600 ms\n5 events")) +
  geom_violin(alpha = 0.5, show.legend = FALSE) +
  geom_boxplot(width = 0.05, show.legend = FALSE, alpha = 0.2) +
  theme_jelle() +
  labs(x = "Condition", y = "G measure", title = "Grammatical redundancy by condition")
ggsave(here("plots", "g-score.pdf"))
```
```



### Tempo difference

```{r tempo-diff}
pp.measures.std <- ddply(pp.measures, .(pp_id, stim_tempo_intended, length), summarize,
  y = tempo_diff_avg,
  ymin = tempo_diff_avg - tempo_diff_sd,
  ymax = tempo_diff_avg + tempo_diff_sd
)


ggplot(pp.measures.std, aes(x = 1, y = y, colour = factor(pp_id), label = pp_id)) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax), position = position_dodge(width = 0.1)) +
  theme_jelle() +
  facet_grid(rows = vars(stim_tempo_intended), cols = vars(length), labeller = labels) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  labs(
    y = "Tempo difference [ms]", title = "Tempo difference by participant and condition",
    subtitle = "Negative values mean participants decreased tempo.", colour = "Participant",
    caption = "Error bars indicate ± SD.\nNote that the values presented here are averages of 
    positive and negative numbers, which complicates interpretation."
  ) +
  theme(legend.title = element_text(size = 12), legend.text = element_text(size = 12))

ggsave(here("plots", "tempo-diff.pdf"))
```
```


### Entropy

```{r entropy}
ggplot(ITIs.bytrial, aes(y = entropy_diff, x = 1, colour = pp_id)) +
  stat_summary(geom = "pointrange", position = position_dodge(width = 0.1)) +
  facet_grid(rows = vars(stim_tempo_intended), cols = vars(length), labeller = labels) +
  theme_jelle() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  labs(
    y = "Entropy difference", title = "Entropy difference by participant and condition",
    subtitle = "Negative values mean participants decreased entropy", colour = "Participant"
  )
ggsave(here("plots", "entropy.pdf"))
```
```
```


### Preference for binary or ternary ratios

```{r binary-ternary-dataprep}
binary.ternary <- reshape2::melt(pp.measures,
  id.vars = c("pp_id", "stim_tempo_intended", "length"),
  measure.vars = c("preference_binary_D", "preference_ternary_D")
)

colnames(binary.ternary) <- c("pp_id", "stim_tempo_intended", "length", "ratio_type", "KS_D")
binary.ternary$ratio_type <- as.character(binary.ternary$ratio_type)
binary.ternary$ratio_type[binary.ternary$ratio_type == "preference_binary_D"] <- "Binary"
binary.ternary$ratio_type[binary.ternary$ratio_type == "preference_ternary_D"] <- "Ternary"
binary.ternary$ratio_type <- as.factor(binary.ternary$ratio_type)

```

```{r binary-ternary}

ggplot(binary.ternary, aes(x = ratio_type, y = KS_D, fill = ratio_type)) +
  geom_violin(alpha = 0.5, show.legend = FALSE) +
  geom_boxplot(width = 0.05, show.legend = FALSE, alpha = 0.2) +
  geom_point(size = 1, show.legend = FALSE, position = position_jitter(width = 0.2)) +
  facet_grid(rows = vars(stim_tempo_intended), cols = vars(length), labeller = labels) +
  theme_jelle() +
  labs(
    x = "Ratio type", y = "Kolmigorov-Smirnov D", title = "Preference for binary vs. ternary ratios",
    subtitle = "Higher values mean more preference", colour = "Participant"
  )
ggsave(here("plots", "binary-ternary.pdf"))
```
```
```

```


## Distributions

### ITI ridges

```{r ITI-ridges}
ggplot(ITIs, aes(x = resp_iti, y = pp_id, fill = pp_id)) +
  geom_density_ridges(alpha = 0.75, show.legend = FALSE) +
  theme_ridges(center = TRUE) +
  facet_grid(rows = vars(stim_tempo_intended), cols = vars(length), labeller = labels) +
  theme_jelle() +
  labs(x = "ITI [ms]", y = "Participant")
ggsave(here("plots", "ITI-ridges.pdf"))
```



### Interval ratios

Here with the conditions plotted as facets:

```{r intervalratios-ridges}
ggplot(interval.ratios, aes(x = interval_ratio, y = factor(pp_id), fill = factor(pp_id))) +
  geom_density_ridges(alpha = 0.75, show.legend = FALSE) +
  theme_ridges(center = TRUE) +
  facet_rep_grid(
    rows = vars(stim_tempo_intended), cols = vars(length), repeat.tick.labels = "x",
    labeller = labels
  ) +
  theme_jelle() +
  xlim(0, 1) +
  scale_x_continuous(labels = c("", "1:3", "1:2", "1:1", "2:1", "3:1", ""), breaks = c(0, 0.25, 0.33, 0.5, 0.66, 0.75, 1)) +
  geom_vline(xintercept = c(0.25, 0.33, 0.5, 0.66, 0.75), linetype = "dashed", color = "grey") +
  labs(x = "Interval ratio", y = "Participant")

ggsave(here("plots", "intervalratios-ridges.pdf"))
```


```{r intervalratios-ridges-condition}
ggplot(interval.ratios, aes(x = interval_ratio, y = factor(pp_id), fill = condition)) +
  geom_density_ridges(alpha = 0.6) +
  # theme_ridges() +
  xlim(0, 1) +
  scale_x_continuous(labels = c("", "1:3", "1:2", "1:1", "2:1", "3:1", ""), breaks = c(0, 0.25, 0.33, 0.5, 0.66, 0.75, 1)) +
  geom_vline(xintercept = c(0.25, 0.33, 0.5, 0.66, 0.75), linetype = "dashed", color = "grey") +
  labs(x = "Interval ratio", y = "Participant", fill = "Condition") +
  theme_jelle(labels = c("400 ms; 4 events", "400 ms; 5 events", "600 ms; 4 events", "600 ms; 5 events"))

ggsave(here("plots", "intervalratios-ridges-2.pdf"))
```