---
title: "NPOR analysis: Behavioural measures"
author: Massimo Lumaca, Jelle van der Werff
date: "`r Sys.Date()`"
output:
  rmdformats::material:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
    cards: true
    code_folding: hide
---

# Preliminaries

## Load packages and data

```{r load-packages, echo = FALSE, warning = FALSE}
# for easy loading of packages
if (!require("pacman")) install.packages("pacman")

# load packages
p_load("here", "ggplot2", "dplyr", "nlme", "boot", "effectsize", "rmdformats", "knitr", "kableExtra")

here <- here::here

# for relative file paths (e.g. here('Input', 'file.csv'))
here::i_am(file.path("stats", "behavioural_results.Rmd"))

# load custom ggplot theme
source(here("scripts", "jelle_ggplot_theme.R"))

# Load ata
pp.measures <- read.csv(here("data", "experiment", "processed", "pp_measures.csv"))
itis <- read.csv(here("data", "experiment", "processed", "ITIs.csv"))
itis.bytrial <- read.csv(here("data", "experiment", "processed", "ITIs_bytrial.csv"))

# set knitr options
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, out.width = "100%", fig.height = 10, fig.width = 10, comment = NA)
```

# Stimulus vs. response
We compare stimulus and response measures using one-sample Wilcoxon signed-rank tests.

```{r only-6005}
pp.measures <- pp.measures[pp.measures$condition == "600_5", ]

```

```{r sttats-and-effect-size-stim-to-resp}

# Entropy
entropy.wc <- wilcox.test(pp.measures$entropy_diff_norm_q, alternative = "two.sided")
entropy.wc
entropy.r <- entropy.wc$statistic[[1]] / sqrt(length(pp.measures$entropy_diff_norm_q))
entropy.r

# Grammatical redundancy
grammatical_redundancy.wc <- wilcox.test(pp.measures$G_resp, alternative = "greater")
grammatical_redundancy.wc
grammatical_redundancy.r <- grammatical_redundancy.wc$statistic[[1]] / sqrt(length(pp.measures$G_resp))
grammatical_redundancy.r

# isochrony introduced
isochrony_introduced.wc <- wilcox.test(pp.measures$isochrony_introduced, alternative = "two.sided")
isochrony_introduced.wc
isochrony_introduced.r <- isochrony_introduced.wc$statistic[[1]] / sqrt(length(pp.measures$isochrony_introduced))
isochrony_introduced.r

# binary/ternary introduced
binary_ternary_introduced.wc <- wilcox.test(pp.measures$binary_or_ternary_introduced, alternative = "two.sided")
binary_ternary_introduced.wc
binary_ternary_introduced.r <- binary_ternary_introduced.wc$statistic[[1]] / sqrt(length(pp.measures$binary_or_ternary_introduced))
binary_ternary_introduced.r

```

# Correlations among measures

```{r correlations}
print(paste("Correlation between binary/ternary and edit distance:", cor(pp.measures$binary_or_ternary_introduced, pp.measures$edit_distance_norm_q)))
# get p value from correlation
cor.test(pp.measures$binary_or_ternary_introduced, pp.measures$edit_distance_norm_q)
print(paste("Correlation between G and edit distance:", cor(pp.measures$G_resp, pp.measures$edit_distance_norm_q)))
cor.test(pp.measures$G_resp, pp.measures$edit_distance_norm_q)
```

# Between-individual variability for all measures

```{r cov-all-measures}

# Entropy
entropy.cov <- sd(pp.measures$entropy_diff_norm_q) / mean(abs(pp.measures$entropy_diff_norm_q))
print(paste("Entropy CV: ", entropy.cov))

# isochrony introduced
isochrony_introduced.cov <- sd(pp.measures$isochrony_introduced) / mean(abs(pp.measures$isochrony_introduced))
print(paste("Isochrony introduced CV: ", isochrony_introduced.cov))

# binary/ternary introduced
binary_ternary_introduced.cov <- sd(pp.measures$binary_or_ternary_introduced) / mean(abs(pp.measures$binary_or_ternary_introduced))
print(paste("Binary/ternary introduced CV: ", binary_ternary_introduced.cov))

# Grammatical redundancy
G_resp.cov <- sd(pp.measures$G_resp) / mean(abs(pp.measures$G_resp))
print(paste("G CV: ", G_resp.cov))


```

  CV for measures range from 0.76 to 1.22.

# Variability between-subjects vs. within-subjects

## Entropy
```{r variability-between-within-entropy}
# start with baseline
entropy.baseline <- gls(entropy_diff_norm_q_avg ~ 1, method = "ML", data = pp.measures)

# add in random factor
entropy.random <- lme(entropy_diff_norm_q_avg ~ 1,
  random = ~ 1 | pp_id_behav,
  method = "ML",
  data = pp.measures
)

# compare models
entropy.model.comp <- as.data.frame(anova(entropy.baseline, entropy.random))

kable(entropy.model.comp, format = "html", caption = "Entropy: between- vs. within-subject variability") %>%
  kable_styling("striped", full_width = FALSE, position = "center")

```

## Binary/ternary ratios

```{r variability-between-within-binaryternary}
# start with baseline
binary_ternary.baseline <- gls(binary_or_ternary_introduced ~ 1, method = "ML", data = pp.measures)

# add in random factor
binary_ternary.random <- lme(binary_or_ternary_introduced ~ 1,
  random = ~ 1 | pp_id_behav,
  method = "ML",
  data = pp.measures
)

# compare models
binary_ternary.model.comp <- as.data.frame(anova(binary_ternary.baseline, binary_ternary.random))

kable(binary_ternary.model.comp, format = "html", caption = "Binary/ternary ratios: between- vs. within-subject variability") %>%
  kable_styling("striped", full_width = FALSE, position = "center")

```

## Isochrony introduced

```{r isochrony-introduced-isoc-introduced}
# start with baseline
isochrony.baseline <- gls(isochrony_introduced ~ 1, method = "ML", data = pp.measures)

# add in random factor
isochrony.random <- lme(isochrony_introduced ~ 1,
  random = ~ 1 | pp_id_behav,
  method = "ML",
  data = pp.measures
)

# compare models
isochrony.model.comp <- as.data.frame(anova(isochrony.baseline, isochrony.random))

kable(isochrony.model.comp, format = "html", caption = "Isochrony introduced: between- vs. within-subject variability") %>%
  kable_styling("striped", full_width = FALSE, position = "center")

```

### Grammatical redundancy

G is already a summary statistic at the level of the individual within condition. As such, we can't test between vs. within-subject variability.
